\section{Introduction}

You wake up.  Your head throbs, and your clothes drip with saltwater. 
Around you lie a few pieces of driftwood, the sight of which triggers
a series of memories: a ship-journey from your homeland, full of promise;
storm clouds on the horizon, roiling and swirling; and the screams of the
sailors on the deck above as the storm caught the ship in its fury, tearing
it apart without mercy.

Somehow, you survived, clinging to several planks that miraculously
remained joined to form a small raft. And just as lucky, you still had all
your possessions, your backpack balanced precariously at the end of the
boards.  Drifting over the cold sea, you ate sparingly, and drank even
less.  Eventually, your water ran out, and, fighting a fever, you drifted
into unconsciousness, expecting to awake on the other side.

But it was not to be.  You stand on the rocky beaches of an island,
your raft shattered against the rocks.  In the distance is some sort of
settlement - where you go from here is entirely up to you.
 
You left in search of adventure: monsters were returning, creatures of myth
and legend becoming startlingly real.  Your friends set sail a year
earlier, carrying their possessions on their backs, seeking their fame and
fortune.  You said your goodbyes, certain that, for one reason or another,
you were unlikely to see them again.

After that year, the wanderlust took you, too.  You took what you could
carry, selling the rest.  A small sailing ship offered passage for a 
handful of ivories, so long as you stayed out of the way.  You
jumped at the opportunity, and now you stand, dripping and exhausted, on a
lonely beach, just barely alive.

The world is huge, and this is but an island.  But first things first:
you need to find a boat.

